\LARGE \bf Energy-Aware Dynamic Planning Algorithm for Autonomous UAVs

### 2020-12-15 icra-2021.tex
implementing Hector's and Ulrik's suggestions (2)

\begin{defn}[Stage, plan, and triggering point]\label{def:mission}
  A \emph{stage} $\Gamma_i$ at time instant $k$ of a plan $\Gamma$ is defined as the ordered list
  \begin{equation}\label{eq:mission}\begin{split}
    \Gamma_i:=\{\langle&\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho}),c_{i,\rho+1},\dots,c_{i,\rho+\sigma}\rangle\\\mid
    &\,\exists\,\,\mathbf{p}_k,\,\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathbb{C}_i,\,\\
      &\,\forall j\in[\sigma]^+,\,c_{i,\rho+j}\in\mathbb{S}_{i,j}\,\},
  \end{split}\end{equation}
  where $\mathbb{C}_i:=[\underline{c}_i,\overline{c}_i]\subseteq\mathbb{R}$ is the trajectory constraints set, and $\mathbb{S}_{i,j}:=[\underline{c}_{i,\rho+j},\overline{c}_{i,\rho+j}]\subseteq\mathbb{Z}_{\geq 0}$ the $j$-th computation constraints set. $\mathbf{p}_k$ is a point of a UAV flying at an assigned altitude $h\in\mathbb{R}_{>0}$ w.r.t. some inertial navigation frame $\mathcal{O}_W$.

  The \emph{plan} is a finite state machine $\Gamma$ where the state-transition function $\delta:\bigcup_i{\Gamma_i}\times\mathbb{R}^2\rightarrow\bigcup_i{\Gamma_i}$ maps a stage and a point to the next stage
  \begin{equation*}\delta(\Gamma_i,\mathbf{p}_k):=\begin{cases}
    \Gamma_{i+1} & \text{if }\mathbf{p}_k=\mathbf{p}_{\Gamma_i}\\
    \Gamma_i & \text{otherwise}
  \end{cases}.\end{equation*}
  The point $\mathbf{p}_{\Gamma_{i}}$ that allows the transition between $\Gamma_i$ and $\Gamma_{i+1}$ is called \emph{triggering point}.
\end{defn}

@@

A slice of the plan in Figure~\ref{fig:state-machine} shows the transition between the stages. The triggering point $\mathbf{p}_{\Gamma_{i-1}}$ allows the transition to stage $\Gamma_i$. The UAV remains in the stage with any generic point $\mathbf{p}_{k_1}$ and $\mathbf{p}_{k_2}$. It eventually enters the stage $\Gamma_{i+1}$ with the triggerint point $\mathbf{p}_{\Gamma_i}$ and so on, until it reaches the last triggering point.

@@

\begin{assm}[Plan periodicity]\label{assm:periodic}
  Given an initial plan $\Gamma$ consisting of $l$ stages $\Gamma_i$ s.t. $i\in[l]^+$ and a generic starting point $\mathbf{p}_k$, there exists an integer positive constant $k<l$ such that
  \begin{equation*}
    \varphi_i-\varphi_{k+i}=c_k,
  \end{equation*}
  where $c_1,\dots,c_k$ are arbitrary constant differences.
\end{assm}

Physically, this means that the plan is periodic, with periodicity $k$. From Assumption~\ref{assm:periodic}, follows the next lemma.

\begin{lem}[Rank of the trajectory functions]
  Suppose Assumption~\ref{assm:periodic} holds. Then
  \begin{equation*}
    \mathrm{rank}\begin{bmatrix}
        \varphi_1 & \varphi_{k+1} & \cdots & \varphi_{(l-k)+1} \\ 
        \varphi_2 & \varphi_{k+2} & \cdots & \varphi_{(l-k)+2} \\
        \vdots & \vdots & \ddots & \vdots \\
        \varphi_k & \varphi_{2k} & \cdots & \varphi_l   
      \end{bmatrix}=2.
  \end{equation*}
\end{lem}
\begin{proof} 
  The proof follows directly from Assumption~\ref{assm:periodic}. The column rank of the matrix in the lemma is the dimension of its column space. This denotates the size of the set of all the possible linear combination of the column vectors. We note that there is a limited number (two) of linear combinations of the column vectors being these linearly dependent (the values of the constant differences $c_1,\dots,c_k$ are the same for each $i$). 
\end{proof}

@@

\begin{pb}[UAV planning problem]
  Consider an initial plan $\Gamma$ satisfying Assumption~\ref{assm:periodic}. We are interested in the planning of the path parameters $c_{i,1},\dots,c_{i,\rho}$ and computations parameters $c_{i,\rho+1},\dots,c_{i,\rho+\sigma}$ for the UAV planning problem.
\end{pb}

@@

Let us consider a periodic signal of period $T\in\mathbb{R}_{> 0}$, and a Fourier series of an arbitrary order $r\in\mathbb{Z}_{\geq 0}$ for the purpose of modeling of the signal
\begin{equation}\label{eq:fourier}
  h(t)=a_0/T+(2/T)\sum_{j=1}^{r}{\left(a_j\cos{\omega jt}+b_j\sin{\omega jt}\right)},
\end{equation}
where $h:\mathbb{R}_{\geq 0}\rightarrow\mathbb{R}$ maps time to the instantaneous energy consumption, $\omega:=2\pi/T$ is the angular frequency, and $a,b\in\mathbb{R}$ the Fourier series coefficients.

Let us consider a linear time-invariant (LTI) state-space model
\begin{equation}\label{eq:state-perf}\begin{split}
  \dot{\mathbf{q}}&=A\mathbf{q}+B\mathbf{u}+\mathbf{w},\\
  y&=C\mathbf{q}+v,
\end{split}\end{equation}
where $y\in\mathbb{R}_{\geq 0}$ is the instantaneous energy consumption, and $\mathbf{w}\in\mathbb{R}^m,v\in\mathbb{R}$ are the contributions of the uncertainty.

@@

\begin{lem}[Signals equality]\label{lem:eqv}Suppose there is no state and output uncertainty (the contributions $\mathbf{w},v$ are zeros) and the control is zero. Suppose further matrices $A,C$ are described by Equation~(\ref{eq:state-details}), and the initial guess $\mathbf{q}_0$ is $\mathbf{q}_0=\begin{bmatrix}a_0 & a_1/2 & b_1/2 & \cdots & a_r/2 & b_r/2\end{bmatrix}$. Then, the signal $h$ in Equation~(\ref{eq:fourier}) is equal to the signal $y$ in Equation~(\ref{eq:state-perf}).
\end{lem}
\begin{proof}
The equivalence of the models is trivial and the equality of the two signals is achieved by a proper choice of items of matrices $A,C$ and the initial guess $\mathbf{q}_0$. We refer the reader to Appendix~\ref{app:proof-eqv} for a formal proof of the equality where we justify the choices of the items of the matrices and of the initial guess. 
\end{proof}

@@

Given $l$ stages $\mathcal{M}_i$ (TEE $\varphi_i$, tasks $\Psi$, computations $\mathbf{s}_i\in\mathbb{S}_i$, and adaptations $\mathbf{c}_i\in\mathbb{C}_i$ for all $i\in[l]$), the main purpose of the algorithm is to output a control input sequence $\mathbf{u}^a:=\{\mathbf{u}_0^a,\mathbf{u}_1^a,\dots\}$ in a valid mission.

\begin{defn}[Valid mission]\label{def:valid}
  A mission is valid if for every stage $\mathcal{M}_{i-1},\,i\in[l]^+$ there exist a control input $\mathbf{u}_k^{a}$ that produce the next stage $\mathcal{M}_i$
  \begin{equation}\begin{split}\label{eq:mission-valid}
    \mathbf{u}^a_{k}=\{&\langle\mathbf{c}_{k},\mathbf{s}_{k}\rangle\mid\,\exists n\in\mathbb{Z}_{>0},\\&\langle\varphi_{i-1}(\mathbf{p}_{k-n},\mathbf{c}_{k-n}),\Psi(\mathbf{s}_{k-n})\rangle\in\mathcal{M}_{i-1}\\
    &\Longrightarrow\langle\varphi_i(\mathbf{p}_{k},\mathbf{c}_{k}),\Psi(\mathbf{s}_{k})\rangle\in\mathcal{M}_{i}\}.
  \end{split}\end{equation}
\end{defn}

>>>>

\begin{defn}[Stage, plan, triggering, and final point]\label{def:mission}
  A \emph{stage} $\Gamma_i$ at time instant $k$ of a plan $\Gamma$ is defined as the ordered list
  \begin{equation}\label{eq:mission}\begin{split}
    \Gamma_i:=\{\langle&\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho}),c_{i,\rho+1},\dots,c_{i,\rho+\sigma}\rangle\\\mid
    &\,\exists\,\,\mathbf{p}_k,\,\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathbb{C}_i,\,\\
      &\,\forall j\in[\sigma]^+,\,c_{i,\rho+j}\in\mathbb{S}_{i,j}\,\},
  \end{split}\end{equation}
  where $\mathbb{C}_i:=[\underline{c}_i,\overline{c}_i]\subseteq\mathbb{R}$ is the trajectory constraints set, and $\mathbb{S}_{i,j}:=[\underline{c}_{i,\rho+j},\overline{c}_{i,\rho+j}]\subseteq\mathbb{Z}_{\geq 0}$ the $j$-th computation constraints set. $\mathbf{p}_k$ is a point of a UAV flying at an assigned altitude $h\in\mathbb{R}_{>0}$ w.r.t. some inertial navigation frame $\mathcal{O}_W$.

  The \emph{plan} is a finite state machine $\Gamma$ where the state-transition function $\delta:\bigcup_i{\Gamma_i}\times\mathbb{R}^2\rightarrow\bigcup_i{\Gamma_i}$ maps a stage and a point to the next stage
  \begin{equation*}\delta(\Gamma_i,\mathbf{p}_k):=\begin{cases}
    \Gamma_{i+1} & \text{if }\mathbf{p}_k=\mathbf{p}_{\Gamma_i}\\
    \Gamma_i & \text{otherwise}
  \end{cases}.\end{equation*}
  The point $\mathbf{p}_{\Gamma_{i}}$ that allows the transition between $\Gamma_i$ and $\Gamma_{i+1}$ is called \emph{triggering point}. The last triggering point $\mathbf{p}_{\Gamma_{l}}$ relative to last stage $\Gamma_l$ is called \emph{final point}.
\end{defn}

@@

A slice of the plan in Figure~\ref{fig:state-machine} shows the transition between the stages. The triggering point $\mathbf{p}_{\Gamma_{i-1}}$ allows the transition to stage $\Gamma_i$. The UAV remains in the stage with any generic point $\mathbf{p}_{k_1}$ and $\mathbf{p}_{k_2}$. It eventually enters the stage $\Gamma_{i+1}$ with the triggering point $\mathbf{p}_{\Gamma_i}$ and so on, until it reaches the last triggering point.

@@

\begin{assm}[Plan periodicity]\label{assm:periodic}
  Given an initial plan $\Gamma$ consisting of $l$ stages, a generic starting point $\mathbf{p}$ and current levels of parameters, there exists a constant $n\in\mathbb{Z}_{>0}, n<l$ such that
  \begin{equation*}
    \varphi_i-\varphi_{n+i}=d_i,\,\,\,\forall i\in[l]^+
  \end{equation*}
  where $d_1,\dots,d_l$ are arbitrary constant differences.
\end{assm}

Physically, this means that the plan is periodic.

\begin{defn}[Period]\label{def:period}
  The period $T\in\mathbb{R}_{> 0}$ is defined as the time between $\varphi_i$ and $\varphi_{n+i}$ for an arbitrary $i\in[l]^+$.
\end{defn}

The algorithm measures the time between the stages $\Gamma_i$ and $\Gamma_{n+i}$ and assumes the initial period is one.

From Assumption~\ref{assm:periodic}, follows the next proposition.

\begin{prop}[Rank of the trajectory functions]\label{prop:rank}
  Suppose Assumption~\ref{assm:periodic} holds. Then
  \begin{equation*}
    \mathrm{rank}\begin{bmatrix}
        \varphi_1 & \varphi_{n+1} & \cdots & \varphi_{(l-n)+1} \\ 
        \varphi_2 & \varphi_{n+2} & \cdots & \varphi_{(l-n)+2} \\
        \vdots & \vdots & \ddots & \vdots \\
        \varphi_n & \varphi_{2n} & \cdots & \varphi_l   
      \end{bmatrix}=2.
  \end{equation*}
\end{prop}
\begin{proof} 
  The proof follows directly from Assumption~\ref{assm:periodic}. The column rank of the matrix in the lemma is the dimension of its column space. It denotes the size of the set of all the possible linear combination of the column vectors. We note that there is a limited number (two) of linear combinations of the column vectors being these linearly dependent (the values of the constant differences $d_1,\dots,d_l$ are the same for each $i$). 
\end{proof}

@@

\begin{pb}[UAV planning problem]\label{pb}
  Consider an initial plan $\Gamma$ defined in Definition~\ref{def:mission} satisfying Assumption~\ref{assm:periodic} and Proposition~\ref{prop:rank}. We are interested in the planning of the path parameters $c_{i,1},\dots,c_{i,\rho}$ and computations parameters $c_{i,\rho+1},\dots,c_{i,\rho+\sigma}$ for the UAV planning problem.
\end{pb}

@@

Let us consider a periodic signal of period $T$, and a Fourier series of an arbitrary order $r\in\mathbb{Z}_{\geq 0}$ for the purpose of modeling of the signal
\begin{equation}\label{eq:fourier}
  h(t)=a_0/T+(2/T)\sum_{j=1}^{r}{\left(a_j\cos{\omega jt}+b_j\sin{\omega jt}\right)},
\end{equation}
where $h:\mathbb{R}_{\geq 0}\rightarrow\mathbb{R}$ maps time to the instantaneous energy consumption, $\omega:=2\pi/T$ is the angular frequency, and $a,b\in\mathbb{R}$ the Fourier series coefficients.

The energy signal can be modeled by Equation~\ref{eq:fourier} and the output of a linear model
\begin{equation}\label{eq:state-perf}\begin{split}
  \dot{\mathbf{q}}&=A\mathbf{q}+B\mathbf{u},\\
  y&=C\mathbf{q},
\end{split}\end{equation}
where $y\in\mathbb{R}$ is the instantaneous energy consumption.

@@

Linear model in Equation~\ref{eq:state-perf} allows us to include the control in the model of Equation~\ref{eq:fourier}. The algorithm uses the model to find the control sequence that results in the optimal parameters configuration (using MPC).

\begin{lem}[Signals equality]\label{lem:eqv}Suppose control is zero, matrices $A,C$ are described by Equation~(\ref{eq:state-details}), and the initial guess $\mathbf{q}_0$ is 
  \begin{equation*}
  \mathbf{q}_0=\begin{bmatrix}a_0 & a_1/2 & b_1/2 & \cdots & a_r/2 & b_r/2\end{bmatrix}^T.
  \end{equation*} 
  Then, the signal $h$ in Equation~(\ref{eq:fourier}) is equal to the signal $y$ in Equation~(\ref{eq:state-perf}).
\end{lem}
\begin{proof}
The equivalence of the models is trivial and the equality of the two signals is achieved by a proper choice of items of matrices $A,C$ and the initial guess $\mathbf{q}_0$. We refer the reader to Appendix~\ref{app:proof-eqv} for a formal proof of the equality where we justify the choices of the items of the matrices and of the initial guess. 
\end{proof}

@@

The main purpose of the algorithm is to output a valid control sequence $\mathbf{u}:=\{\mathbf{u}_0,\mathbf{u}_1,\dots\}$ at each time step given an initial plan $\Gamma$, to solve Problem~\ref{pb}.

----

\begin{defn}[Valid control sequence]\label{def:valid}
  A control sequence $\mathbf{u}$ is valid if for every stage $\Gamma_{i-1},\,i\in[l]^+$ if there exist a control $\mathbf{u}_k$ that produce the next stage $\Gamma_i$
  \begin{equation}\begin{split}\label{eq:mission-valid}
    \mathbf{u}^a_{k}=\{&\langle\mathbf{c}_{k},\mathbf{s}_{k}\rangle\mid\,\exists n\in\mathbb{Z}_{>0},\\&\langle\varphi_{i-1}(\mathbf{p}_{k-n},\mathbf{c}_{k-n}),\Psi(\mathbf{s}_{k-n})\rangle\in\mathcal{M}_{i-1}\\
    &\Longrightarrow\langle\varphi_i(\mathbf{p}_{k},\mathbf{c}_{k}),\Psi(\mathbf{s}_{k})\rangle\in\mathcal{M}_{i}\}.
  \end{split}\end{equation}
\end{defn}
